---
title: '4'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
Mdata<-read.csv("Pharmaceuticals.csv")
head(Mdata)

dim(Mdata)  # In this data frame there are 14 variables and 21 records  

str(Mdata) # It Displays the number of variables as well as their data types.

U<-unique(Mdata)  # This indicates that there are no duplicate rows
dim(U)

pairs(Mdata[3:11]) # plotting all possible discrete variables

library(DataExplorer)
introduce(Mdata) # It shows the total number of discrete and continuous variables,as well as any missing values


# There are 9 Continuous variables , 5 discrete variables and 0 missing variables 
```

# Data Exploration
```{r}
set.seed(123)
plot(Market_Cap~Rev_Growth,Mdata)
with(Mdata, text(Market_Cap~Rev_Growth, labels=Symbol,pos=4,cex=0.4))    # A general estimation of how the market cap vs revenue growth distribution looks.

plot(Market_Cap~Net_Profit_Margin,Mdata)
with(Mdata, text(Market_Cap~Net_Profit_Margin,labels=Symbol,pos=4,cex=0.4))
```

K-Means Clustering Model

```{r}
library(tidyverse)
library(factoextra)
library(ISLR)


Mdata1<-scale(Mdata[, c(3:11)], center=TRUE,scale=TRUE)

dist_Mdata<-get_dist(Mdata1)
print(dist_Mdata,digits=3)

heatmap(Mdata1,scale="column")
#Companies 8 and 17 are the most distant  from one another,with a distance of 7.18 miles

Mdata[c(8,17),] # From this we  can observe that there is a significant difference in market capital,revenue growth and net profit profit margin between these two organizations.


boxplot(Mdata1,col=c("deepskyblue","deepskyblue4","cadetblue","cadetblue1","cadetblue2","cadetblue3","cadetblue4","dodgerblue","dodgerblue4"), las=2)
```

Finding optimal number for k as per Elbow method

```{r}
fviz_nbclust(Mdata1,kmeans,method="wss")+
  geom_vline(xintercept=2,linetype=2)   #Elbow Method
fviz_nbclust(Mdata1,kmeans,method= "silhouette")  #Silhouette Method
```
# Outliers have a significant impact on the silhouette approach of determining the appropriate amount of k.Because our data has a large number of outliers.

# using the silhouette method to remove the outliers,the ideal value of k was likewise found to be 2.

# Hence,we can consider the optimal number of k using the Elbow method.
 
```{r}
#k=2
k_mean_model2<-kmeans(Mdata1,centers=2,nstart=25)
fviz_cluster(k_mean_model2,data=Mdata1)+labs(subtitle="k=2")
k_mean_model2
```
 
# b. Interpret the clusters with respect to the numerical variables used in forming the clusters.

Companies in cluster 2 have a higher Return on Equity than companies in cluster 1 for the identical Return on Equity values.

# c. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters) 
 
Cluster 1 has highest rev_growth, highest net_profit_margin and has hold median recommendation.

Cluster 2 has low rev_growth, high net_profit_margin and has moderate buy recommendation.
 
# d. Provide an appropriate name for each cluster using any or all of the variables in the dataset.

Large size company : cluster 1

Small size company : cluster 2















